Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-15 paper/sec-mcsamples.tex
Sect 2.3

\subsection{Higgs pair production}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%
Parton level events are then showered with the {\tt Pythia8} Monte
Carlo~\cite{Sjostrand:2007gs,Sjostrand:2014zea}, in particular with {\tt v8.201}.
%
We use the default settings in the parton shower, in particular
we use the Monash 2013 tune~\cite{Skands:2014pea}, based on the NNPDF2.3LO PDF set~\cite{Ball:2012cx}.
%
We have switched off hadronisation, since this allows a more transparent modeling of
$b$-tagging.
%
A complete simulation of $b$-tagging at the hadron level is difficult to achieve without a full
detector simulator, which is beyond the scope of this paper.
%
For a similar reason, we don't consider underlying event, multiple parton interactions and
pile-up, which can be taken of using a variety of subtraction methods~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba}.
%
The same applies to the background samples, which we discuss now.

@@

Finally, let us discuss the smearing applied to signal
and background samples.
%
It is beyond the scope of this work to perform a complete study of the impact of
detector effects in this process.
%
However, it is still important to include some form of rough estimate of detector
effects in the analysis, to properly estimate for example the impact
of the Higgs mass window cuts.
%
In this work we simulate the finite momentum resolution of the ATLAS and CMS
hadronic calorimeters by applying a Gaussian smearing to the four components
of the four-momentum $p$, with mean zero and standard deviation $\Delta p$.
%
We take as baseline value a smearing of $\Delta p=5\%$, and in
Sect.~\ref{sec:optimisation} we explore how our results change
if a more optimistic or pessimistic value is assumed for the momentum
resolution in the LHC detectors.

>>>>

\subsection{Higgs pair production by gluon-fusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%
Parton level events are then showered with the {\tt Pythia8} Monte
Carlo~\cite{Sjostrand:2007gs,Sjostrand:2014zea}, in particular with {\tt v8.201}.
%
We use the default settings in the parton shower, in particular
we use the Monash 2013 tune~\cite{Skands:2014pea}, based on the NNPDF2.3LO PDF set~\cite{Ball:2012cx}.
%
We have switched off hadronisation, since this allows a more transparent modeling of
$b$-tagging.
%
A complete simulation of $b$-tagging at the hadron level is difficult to achieve without a full
detector simulator, which is beyond the scope of this paper.

@@

\subsection{Mimicking detector resolution}

While it is beyond the scope of this work to perform a full
detector simulation, it is still important to include an estimate of detector
effects in the analysis, in particular for the finite resolution
in energy and momentum, which will affect some kinematical variables, in particular
the invariant mass of the Higgs candidates.
%
For example, in the absence of any four-momentum smearing, the impact of the $m_h$
distributions of the Higgs candidates in the MVA
would be unrealistically large.

In this work we simulate the finite momentum resolution of the ATLAS and CMS
hadronic calorimeters by applying a Gaussian smearing
with mean zero and standard deviation $\sigma_E$.
%
In particular, we rescale the four-momentum of all
final-state particles, before jet clustering, using
a common factor for the energy $E$ and the length of the
three-momentum $\vec{p}$, while keeping the direction unchanged:
%
\be
E_i \, \to \, E'_i= \lp 1+ r_i\cdot\sigma_E \rp\, E_i \, ,\quad
\vec{p}_i \, \to \, \vec{p}'_i= \lp 1+ r_i\cdot\sigma_E \rp\, \vec{p}_i \, , \quad
i=1,\ldots,N_{\rm part} \, ,
\ee
with $r_i$ an univariate Gaussian random number which is different for each
of the $N_{\rm part}$ particles in the event.

This smearing reflects the fact that the angular resolution of the LHC detectors
is better than the energy resolution.
%
We take as baseline value a momentum smearing
factor of $\sigma_E=5\%$, and in
Sect.~\ref{sec:optimisation} we explore how our results change
if a more optimistic or pessimistic value is assumed for this value
of the momentum
resolution in the LHC detectors.

It is also beyond the scope of this paper to
consider
the effects in our analysis of  Underlying Event (UE),
Multiple Parton Interactions (MPI) and in
particular
pile-up (PU).
%
While PU will be be one of the major problems at the HL-LHC, fortunately
a number of powerful methods for the subtraction of  UE, MPI and PU
effects 
have been developed recently~\cite{Cacciari:2009dp,TheATLAScollaboration:2013pia,Butterworth:2008iy,Cacciari:2007fd,Krohn:2009th,Krohn:2013lba,Ellis:2009me,Bertolini:2014bba,Cacciari:2014gra,Cacciari:2014jta,Berta:2014eza,Larkoski:2014wba}.
%
For these reasons, in our Monte Carlo analysis we switch off MPI and PU,
and postpone to a future work a full detector simulation
to quantify the robustness
of our analysis strategy with respect to the inclusion
(and its subtraction) of the large PU
expected at the HL-LHC.

We note that while substructure variables can be affected by PU, event after its
subtraction, our analysis should be relatively robust since we do not cut
over any specific variables, but input the full distributions to the MVA.
%
In particular, if after PU suppression a given variable is modified
but in the same way for signal and background samples, it should not
affect MVA discrimination.
